---
title: "Birdcoin analysis"
output:
  bookdown::html_document2:
    code_folding: hide
    number_sections: true
    fig_caption: true
    toc: true
    toc_float: true
    toc_depth: 3
    theme: lumen
    highlight: tango
bibliography: DEW_reference.bib
link-citation: yes
linestretch: 1.5
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, include=FALSE}
#install.packages("COINr")
library(COINr)
library(targets)
library(dplyr)
library(knitr)
library(kableExtra)
library(purrr)
library(tibble)
library(ggplot2)
library(patchwork)
```

## Read coin

coin_1 from 070_birdcoin.R with all variables for threats and exposure but a subset of variables for sensitivity.

The coin is treated (log transformed for generation length and eoo) and normalised (min-max).

```{r input}

tar_load(coin_1, store = tars$birdcoin$store)
coin_1
plot_framework(coin_1, type = "stack", text_size = 4)

```

## Check percentage of dominance pairs

The greater the number of dominance (or robust) pairs in a classification, the less sensitive the units will be to methodological assumptions.

No dominance pairs found.

```{r dominance_pair}

outlist <- outrankMatrix(coin_1$Data$Normalised[, -c(1:2)])
outlist$nDominant
outlist$fracDominant

```
## Descriptive stats

```{r stats}

stat_table <- get_stats(coin_1, dset = "Normalised", out2 = "df")

kbl(stat_table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "bordered"),
                full_width = FALSE) %>% 
  scroll_box(height = "300px", width = "100%")

```

## Correlations

Finding correlations between indicators that exceed or fall below a given threshold, within aggregation groups.

Weak correlations cause less information to be transferred to the aggregate.

Negative correlations: A negative correlation means that one indicator is capturing a trend that is opposite to the others in the group and may be measuring different features.

```{r correlation}

get_corr_flags(coin_1, dset = "Normalised", cor_thresh = 0.75,
               thresh_type = "high", grouplev = 2)

get_corr_flags(coin_1, dset = "Normalised", cor_thresh = -0.5,
               thresh_type = "low", grouplev = 2)

plot_corr(coin_1, dset = "Normalised", iCodes = list("Threat"),
          grouplev = 2, flagcolours = T)
plot_corr(coin_1, dset = "Normalised", iCodes = list("Sensitivity"),
          grouplev = 2, flagcolours = T)
plot_corr(coin_1, dset = "Normalised", iCodes = list("Exposure"),
          grouplev = 2, flagcolours = T)
```
## Internal consistency reliability

A high c-alpha, or equivalently a high “reliability”, indicates that the individual indicators measure the latent phenomenon well. Nunnally (1978) suggests 0.7 as an acceptable reliability threshold. Yet some authors use .75 or .80 as a cut-off value, while others are as lenient as to go to 0.6. In general this varies by discipline. If the variances of the individual indicators vary widely, as in our test case, a standard practice is to standardise the individual indicators to a standard deviation of 1 before computing the coefficient alpha.

The value of c-alpha increases as the number of indicators increases and as the covariance between each pair of indicators grows.

If the reliability coefficient increases after deleting a specific indicator from the set, it suggests that variable is not highly correlated with the others and may be considered "noise" within that group.

$$
\alpha = 
\frac{k}{k - 1}
\left(
1 - \frac{\sum_{i=1}^{k} \sigma_i^2}{\sigma_T^2}
\right)
$$
where:

- $k$ = number of items (e.g., traits or components in the index)

- $\sigma_i^2$ = variance of item $i$

- $\sigma_T^2$ = variance of the total score 
  (i.e., the variance of $\sum_{i=1}^{k} X_i$)

- $X_i$ = observed score for item $i$

$$
\alpha =
\frac{k \, \bar{c}}
{\bar{v} + (k - 1)\bar{c}}
$$
where:

- $k$ = number of items

- $\bar{c}$ = mean inter-item covariance

- $\bar{v}$ = mean item variance

```{r cronbach}

iMeta <- coin_1$Meta$Ind

# 1. Identify all unique Parent codes at Level 2
parents_lv2 <- iMeta$iCode[iMeta$Level == 2]

# 2. Create a table by looping through parents and calculating c-alpha
ca_table <- data.frame(
  Parent_Group = parents_lv2,
  Cronbach_Alpha = sapply(parents_lv2, function(p) {
    # get_cronbach automatically identifies Level 1 indicators 
    # that belong to parent 'p' when Level = 1 is specified
    get_cronbach(coin_1, dset = "Normalised", iCodes = p, Level = 1)
  })
)

kbl(ca_table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "bordered"),
                full_width = FALSE)

```

## PCA analysis

```{r pca}

pca_sens <- get_PCA(coin_1, dset = "Raw", iCodes = "Sensitivity", out2 = "list")
pca_exp <- get_PCA(coin_1, dset = "Raw", iCodes = "Exposure", out2 = "list")
pca_threat <- get_PCA(coin_1, dset = "Raw", iCodes = "Threat", out2 = "list")

extract_pca_summary <- function(pca_object, dataset_name) {
  
  pca_object$PCAresults %>%
    imap_dfr(function(group_obj, group_name) {
      
      # Get PCA summary matrix
      s <- summary(group_obj$PCAres)$importance
      
      # Convert to dataframe
      as.data.frame(t(s)) %>%
        rownames_to_column("PC") %>%
        mutate(
          dataset = dataset_name,
          group = group_name
        )
      
    })
}

pca_summary_table <- bind_rows(
  extract_pca_summary(pca_sens,   "Sensitivity"),
  extract_pca_summary(pca_exp,    "Exposure"),
  extract_pca_summary(pca_threat, "Threat")
) %>% mutate(across(where(is.numeric), round, 3))

kbl(pca_summary_table) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "bordered"),
                full_width = FALSE) %>% 
  scroll_box(height = "300px", width = "100%")


```

## Check skewness and outlier treatment

```{r treatment}

coin_1$Analysis$Treated$Dets_Table %>%
  arrange(check_SkewKurt0.Pass) %>%   # FALSE first
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "bordered"),
                full_width = FALSE) %>% 
  scroll_box(height = "300px", width = "100%")

plot_dist(coin_1, dset = "Treated", iCodes = "Exposure", Level = 1, type = "Violindot")

```

## Aggregation - equal weight

```{r agg_ew}

coin_1 <- Aggregate(coin_1, dset = "Normalised", f_ag = "a_amean")

plot_corr(coin_1, dset = "Aggregated", iCodes = list("Sensitivity"), withparent = "family", flagcolours = T)
plot_corr(coin_1, dset = "Aggregated", iCodes = list("Exposure"), withparent = "family", flagcolours = T)
plot_corr(coin_1, dset = "Aggregated", iCodes = list("Pressure"), withparent = "family", flagcolours = T)
```

## Visualise the composite indicator and its components

```{r index}

# Vulnerability
plot_bar(coin_1, dset = "Aggregated", 
         iCode = "Vulnerability", 
         uLabel = "uName",
         bar_colours = c("darkslategrey", "lightpink", "royalblue3"),
         stack_children = TRUE) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# Exposure

plot_bar(coin_1, dset = "Aggregated", 
         iCode = "Exposure", 
         uLabel = "uName",
         bar_colours = c("darkslategrey", "lightpink", "royalblue3", 
                         "tan", "olivedrab", "gold", "red4"),
         stack_children = TRUE) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# Sensitivity

plot_bar(coin_1, dset = "Aggregated", 
         iCode = "Sensitivity", 
         uLabel = "uName",
         bar_colours = c("darkslategrey", "lightpink", "royalblue3", "tan", "olivedrab"),
         stack_children = TRUE) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# Pressure

plot_bar(coin_1, dset = "Aggregated", 
         iCode = "Pressure", 
         uLabel = "uName",
         bar_colours = c("darkslategrey", "lightpink", "royalblue3"),
         stack_children = TRUE) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```

## Visulise indicators

```{r vis_ind}

## Sensitivity indicators

sens_iCodes <- na.omit((iMeta$iCode[iMeta$Parent == "Sensitivity"]))
sens_data <- coin_1$Data$Aggregated %>% 
  select(uCode, sens_iCodes) %>% 
  right_join(coin_1$Meta$Unit, by = "uCode")


plots <- lapply(sens_iCodes, function(code) {
  plot_bar(
    coin_1,
    dset = "Aggregated",
    iCode = code,
    uLabel = "uName"
  ) +
    ggtitle(code) +
    theme(
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 4)
    )
})

wrap_plots(plots, ncol = 2)


## Exposure indicators

exp_iCodes <- na.omit((iMeta$iCode[iMeta$Parent == "Exposure"]))

plots <- lapply(exp_iCodes, function(code) {
  plot_bar(
    coin_1,
    dset = "Aggregated",
    iCode = code,
    uLabel = "uName"
  ) +
    ggtitle(code) +
    theme(
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 4)
    )
})

wrap_plots(plots, ncol = 2)

## Pressure indicators

pres_iCodes <- na.omit((iMeta$iCode[iMeta$Parent == "Pressure"]))

plots <- lapply(pres_iCodes, function(code) {
  plot_bar(
    coin_1,
    dset = "Aggregated",
    iCode = code,
    uLabel = "uName"
  ) +
    ggtitle(code) +
    theme(
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 4)
    )
})

wrap_plots(plots, ncol = 2)
```

## The effect of removing Exposure, Sensitivity, or Pressure

```{r remove_element}

remove_Lv3 <- COINr::remove_elements(coin_1, Level = 3, dset = "Aggregated", iCode = "Vulnerability")

remove_Lv3$RankDiffs

```

## Comparing coins

```{r compare}

tar_load(coinA_1, store = tars$birdcoin$store)
compare_coins(coin_1, coinA_1, 
              dset = "Aggregated", 
              iCode = "Vulnerability",
              compare_by = "ranks",
              sort_by = "Abs.diff", 
              decreasing = TRUE) %>% 
  left_join(coin_1$Meta$Unit, by = "uCode")

```

## Change aggregation method: Aggregation

Compensability: Using the maximum value is a form of non-compensatory aggregation at that level. It means a unit is judged only by its "best" trait, and poor performance in other indicators within that same group is completely ignored.

Information Loss: The sources note that extreme operators result in only one indicator defining the overall performance of that group, which means there is no "balancing" or "optimization" of weights to be performed for that specific level.

```{r aggmax}

# Remove the rest of threat indicators

coin_11 <- change_ind(coin_1, drop = c("t_score_max_sp",
                                      "t_n_lv1",
                                      "t_n_lv2_total"), regen = TRUE)
plot_framework(coin_11, type = "stack", text_size = 4)

# Adjust weight for collisions and level 1 indicator aggregation

w1 <- coin_11$Meta$Weights$Original
w1$Weight[w1$iCode == "CollisionWind"] <- 0.5
w1$Weight[w1$iCode == "CollisionSolar"] <- 0.5
my_max <- function(x, ...) max(x, na.rm = TRUE)

## Lv1-Lv2 max
coin_12 <- Aggregate(coin_11, dset = "Normalised", 
                     w = w1, 
                     f_ag = c("my_max", "a_amean", "a_amean"),
                     regen = TRUE)

## Lv2-lv3 max
coin_13 <- Aggregate(coin_11, dset = "Normalised", 
                     w = w1, 
                     f_ag = c("a_amean", "my_max", "a_amean"),
                     regen = TRUE)

## gmean
coin_14 <- coin_11
coin_14$Log$Normalise$global_specs <- list(f_n = "n_minmax",
                                           f_n_para = list(l_u = c(1,100)))
coin_14 <- Regen(coin_14) %>% 
  Aggregate(dset = "Normalised",
            w = w1, 
            f_ag = "a_gmean")

coin_14$Log
```
## Compare coin_12 to base coin

```{r compare12}

df_12 <- get_results(coin_12, dset = "Aggregated", tab_type = "Aggs")
base_results <- get_results(coin_1, dset = "Aggregated", tab_type = "Aggs")

# Get group with more than 1 indicator
group <- coin_11$Meta$Ind %>%
  group_by(Parent) %>%
  filter(n() > 3, Level == 2) %>%
  ungroup() %>% select(iCode)

cols <- c("Sensitivity", "Exposure", "Pressure")
cols2 <- coin_12$Meta$Ind %>% filter(Level == 2) %>% select(iCode) %>% unique()
  
compare_density_multi(list(base = base_results, coin12 = df_12), cols2$iCode)

# which max indicators were selected?
maxcount <- lapply(cols2$iCode, function(x) {get_max_indicator(coin_12, x) %>% .$summary}) %>%
  purrr::map_dfr(~ tibble(
    iCode = names(.x),
    Percentage = as.numeric(.x)/nrow(coin_12$Data$Raw)
  )) %>% 
  right_join(coin_12$Meta$Ind %>% select(iCode, Parent, iName),
             by = "iCode") %>% 
  mutate(
    Percentage = case_when(
      is.na(Percentage) & Parent %in% cols2 ~ 0,
      is.na(Percentage) & !Parent %in% cols2 ~ 1,
      TRUE ~ round(Percentage, 2)
    )
  )

```

## Compare coin_13 to base coin

```{r compare13}

df_13 <- get_results(coin_13, dset = "Aggregated", tab_type = "Aggs")

cols <- c("Sensitivity", "Exposure", "Pressure")

compare_density_multi(list(base = base_results, coin13 = df_13), 
                      cols) 

# which max indicators were selected?
maxcount <- lapply(cols, function(x) {get_max_indicator(coin_13, x) %>% .$summary}) %>%
  purrr::map_dfr(~ tibble(
    uCode = names(.x),
    MaxCount = as.numeric(.x)
  ))

```
## Compare coin_14 to base coin

```{r compare14}

df_14 <- get_results(coin_14, dset = "Aggregated", tab_type = "Aggs")

cols <- c("Sensitivity", "Exposure", "Pressure")

cols2 <- coin_14$Meta$Ind %>% filter(Level == 2) %>% select(iCode) %>% unique()

compare_density_multi(list(base = base_results, coin14 = df_14), 
                      cols) 

compare_density_multi(list(base = base_results, coin14 = df_14), 
                      cols2$iCode) 

```

## compare coin_12 and 14

```{r compare1214}

compare_density_multi(list(base = base_results, 
                           coin12 = df_12,
                           coin14 = df_14), 
                      cols2$iCode) 

compare_density_multi(list(base = base_results, 
                           coin12 = df_12,
                           coin13 = df_13,
                           coin14 = df_14), 
                      cols) 

compare_coins_multi(list(coin1 = coin_11,
                         coin12 = coin_12,
                         coin13 = coin_13,
                         coin14 = coin_14),
                    dset = "Aggregated",
                    iCode = "Vulnerability",
                    also_get = c("uName"),
                    tabtype = "Values",
                    ibase = 1,
                    sort_table = TRUE,
                    compare_by = "ranks"
)

compare_coins_multi(list(coin1 = coin_11,
                         coin12 = coin_12,
                         coin13 = coin_13,
                         coin14 = coin_14),
                    dset = "Aggregated",
                    iCode = "Vulnerability",
                    also_get = c("uName"),
                    tabtype = "Values",
                    ibase = 1,
                    sort_table = TRUE,
                    compare_by = "scores"
)

```

